---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```


### Download the data
```{r dl}
# Download and unzip the file from assignment web page. Get in the right directory in R.
data<-read.csv("activity.csv")
```

```{r packages}
library(stats)
library(dplyr)
```
## Number of steps per day (ignoring missing values)
```{r stepsPerDay}
par(mfrow=c(1,1))
stepsPerDay<-aggregate(data=data,FUN=sum,steps~date)
hist(stepsPerDay$steps,main="Histogram Steps per day",xlab = "Steps per day",col="red")
meansteps<-mean(stepsPerDay$steps)
mediansteps<-median(stepsPerDay$steps)
```
The mean and median of the total number of steps taken per day are respectively `r meansteps`  and `r mediansteps` .

## What is the average daily activity pattern ?

```{r avg}
stepsInterAvg<-aggregate(data=data,FUN=mean,steps~interval)

par(mfrow=c(1,1))
plot(unique(data$interval),stepsInterAvg$steps,main="Evolution of average steps during a day",xlab="Time",ylab="Number of steps (average)",type="l",col="blue")
maxinterval<-which.max(stepsInterAvg$steps)
maxtime<-stepsInterAvg$interval[maxinterval]

```
The maximum amount of steps done is in the inteval : `r maxtime` .

## Imputing missing values
```{r NAs}
percentNA<-mean(is.na(data))*100
```
The percentage of missing values in the dataset is `r percentNA` % .  
We are going to replace missing values by the average of steps for that 5 minute interval.

```{r meansteps2,mediansteps2}
par(mfrow=c(1,1))
newdata<-data
for (i in 1:length(data$steps)){
    if (is.na(data$steps[i])){
        tmp<-data$interval[i]
        tmp2<-which(stepsInterAvg$interval==tmp)
        newdata$steps[i]<-stepsInterAvg$steps[tmp2]
    }
}
stepsPerDay2<-aggregate(data=newdata,FUN=sum,steps~date)
hist(stepsPerDay2$steps,main="Histogram Steps per day (changed NAs)",xlab = "Steps per day",col="red")
meansteps2<-mean(stepsPerDay2$steps)
mediansteps2<-median(stepsPerDay2$steps)
```
The new mean and median of total number of steps taken per day is respectively `r meansteps2` and `r mediansteps2` .  
These values are slightly under the previous values, it is about 100 steps down.

## Are there differences in activity patterns between weekdays and weekends ?

```{r weekdays, fig.height=8}
newdata["day"]<-character()
newdata$day<-as.character(newdata$day)
newdata$date<-as.Date(newdata$date)
tmp<-weekdays(newdata$date)
for (i in 1:length(newdata$date)){
    if(tmp[i]=="Samedi" || tmp[i]=="Dimanche"){
        newdata$day[i]<-"weekend"
    }
    else {
        newdata$day[i]<-"weekday"
    }
}
newdata$day<-as.factor(newdata$day)
stepsInterAvg2<-aggregate(data=newdata,FUN=mean,steps~interval+day)
par(mfrow=c(2,1))
plot(unique(stepsInterAvg2$interval),stepsInterAvg2$steps[1:288],main="Evolution of avg steps during weekday",xlab="Time",ylab = "Steps",type="l",ylim=c(0,250))

plot(unique(stepsInterAvg2$interval),stepsInterAvg2$steps[289:576],main="Evolution of avg steps during weekend",xlab="Time",ylab = "Steps",type="l",ylim=c(0,250))

```
